---
title: "Refreshing Data from an API in an R Shiny App"
author: "Jesse Harris"
date: "2017-08-29"
output: html_document
---

Posting this in case it helps others like me who are new to Shiny. I was working on a Shiny application with the following requirements:

- When the application loads, pull source data from an API.
- Automatically refresh the data every *x* seconds or minutes thereafter.
- Also allow the user to refresh data manually via an `actionButton`.

I was able to accomplish most of what I needed using these examples:

- ["Reactive Programming with R Shiny" on alshum.com](http://www.alshum.com/shiny-reactive/)
- [reactiveTimer examples on RStudio.com](https://shiny.rstudio.com/reference/shiny/latest/reactiveTimer.html)

The methods I used are shown in the example code below. To keep things simple, I've used a call to `Sys.time()` to simulate an API call that retrieves continuously updated data. In the example, the timer is set to refresh the data every 5 seconds.

```{r eval=FALSE}
library(shiny)
library(lubridate)

ui <- fluidPage(
        actionButton("refreshData","Refresh data"),
        h3("Current seconds. Will increment by five unless manually refreshed."),
        h3(textOutput("currentTime"))
)

server <- function(input, output) {
        
        # initialize reactivevalues
        myReactiveData  <- reactiveValues(myAPIRequest = NULL)
        
        # create a timer
        autoInvalidate  <- reactiveTimer(5000)
        
        observeEvent(
                # events we're looking for
                eventExpr = {
                        # button is pressed
                        input$refreshData
                        # timer runs down
                        autoInvalidate()
                },
                # code to run when one of the above events happens
                handlerExpr = {
                        # where you would put your API call
                        myAPIRequest <- Sys.time()
                        
                        # do some data processing
                        processedResults <- round(second(myAPIRequest))
                        
                        # assign processed data to reactivevalues object
                        myReactiveData$dataToDisplay <- processedResults
                },
                # make handlerExpr run once when app first starts
                ignoreNULL = FALSE
        )
        
        output$currentTime <- renderText(
                myReactiveData$dataToDisplay
        )
}

shinyApp(ui, server)
```

This partially met the requirements, but there was a problem: manually refreshing the data did not affect the timer. If a user manually refreshed the data 3 seconds after initially loading the app, I wanted the timer to reset and wait 5 more seconds before the next automatic refresh. This wasn't happening; the automatic refresh stuck to its original schedule and refreshed every 5 seconds from when the application originally loaded, regardless of user input.

I eventually found [this Stack Overflow article](https://stackoverflow.com/questions/42379402/how-to-terminate-reactivetimer-in-shiny), the key insight from which was the idea of making the timer function itself a reactivevalues object, like so:

```{r eval=FALSE}
autoInvalidate  <- reactiveValues(timer = reactiveTimer(5000))
```

In combination with that, I used a second `observeEvent` to look for the manual refresh and reset the timer. The complete updated code looks like this:

```{r eval=FALSE}
library(shiny)
library(lubridate)

ui <- fluidPage(
        actionButton("refreshData","Refresh data"),
        h3("Current seconds. Will increment by five unless manually refreshed."),
        h3(textOutput("currentTime"))
)

server <- function(input, output) {
        
        # initialize reactivevalues
        myReactiveData  <- reactiveValues(myAPIRequest = NULL)
        autoInvalidate  <- reactiveValues(timer = reactiveTimer(5000))
        
        observeEvent(
                # events we're looking for
                eventExpr = {
                        # button is pressed
                        input$refreshData
                },
                handlerExpr = {
                        # reset the timer
                        autoInvalidate$timer <- reactiveTimer(5000)
                }
        )
        
        observeEvent(
                # events we're looking for
                eventExpr = {
                        # button is pressed
                        input$refreshData
                        # timer runs down
                        autoInvalidate$timer()
                },
                # code to run when one of the above events happens
                handlerExpr = {
                        # where you would put your API call
                        myAPIRequest <- Sys.time()

                        # do some data processing
                        processedResults <- round(second(myAPIRequest))
                        
                        # assign processed data to reactivevalues object
                        myReactiveData$dataToDisplay <- processedResults
                },
                # make handlerExpr run once when app first starts
                ignoreNULL = FALSE
        )
        
        output$currentTime <- renderText(
                myReactiveData$dataToDisplay
        )
}

shinyApp(ui, server)
```

So that's it. In summary, you can do this in Shiny by using multiple `observeEvent` statements and assigning `reactiveTimer` to a reactivevalues object. I don't know if this is the best way to accomplish this, but it worked for me.
